---
title: "Host/Soil-specific Barplots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}

library(tibble)
library(dplyr)
library(ggplot2)

```

### Read in necessary files

```{r}

# count matrices from unrarefied data
unrare_ro_full_mat <-
  readRDS("../intermediate_files/rds/unrare_ro_full_mat.rds")

unrare_ak_full_mat <-
  readRDS("../intermediate_files/rds/unrare_ak_full_mat.rds")

# taxonomy table from full unrarefied physeq object 
full_tax_table_df <- readRDS("../intermediate_files/rds/full_tax_table_df.rds")

# a vector of host names + soil
host_names <- readRDS("../intermediate_files/rds/host_names.rds")

```

## Data Formatting

```{r}

# rename and convert matrices to data frames
ro_barplot_df <- as.data.frame(unrare_ro_full_mat)
ak_barplot_df <- as.data.frame(unrare_ak_full_mat)


saveRDS(ro_barplot_df, "../intermediate_files/rds/ro_barplot_df.rds")
saveRDS(ak_barplot_df, "../intermediate_files/rds/ak_barplot_df.rds")


```

## Prepare data frames for host- and soil-specific ASVs

```{r}

get_unique_asvs <- function (a_host, barplot_df, taxonomy_df) {
  
  # subset host specific ASVs by host. Counts for all other hosts should be 0, 
  # so the rowsum == the host count
  unique_subset <- barplot_df[rowSums(barplot_df)==barplot_df[[a_host]],]
  
  # make asv rownames into a column
  unique_subset <- rownames_to_column(unique_subset, "ASV")
  
  # make a new df with asv names and host sample asv counts as columns
  unique_subset <- select(unique_subset, ASV, a_host)
  
  # take sum of all unique host sample sequences
  total_host_sum <- sum(barplot_df[[a_host]])
  
  # make new relative abundance column by dividing unique count by total
  unique_subset <- mutate(unique_subset, rel_abun=((unique_subset[[a_host]] / total_host_sum) * 100))
  
  # make new column taking family from tax_table 
  # where asv names match in both dfs
  unique_subset$Family <- 
    taxonomy_df$Family[match(unique_subset$ASV, taxonomy_df$ASV)]
  
  # make new column with host name filling all cells
  unique_subset$Host <- rep(a_host, nrow(unique_subset))
  
  # make new column with ASV total per host filling all cells
  unique_subset$Total <- rep(sum(unique_subset[[a_host]]), nrow(unique_subset))
  
  # change host colname for future rbind
  colnames(unique_subset)[2] <- "Count"
  
  return(unique_subset)

}


# run function

ro_unique_asvs <- lapply(host_names, 
                         get_unique_asvs, 
                         ro_barplot_df, 
                         full_tax_table_df)

names(ro_unique_asvs) <- host_names

saveRDS(ro_unique_asvs, "../intermediate_files/rds/ro_unique_asvs.rds")


ak_unique_asvs <- lapply(host_names, 
                         get_unique_asvs, 
                         ak_barplot_df, 
                         full_tax_table_df)

names(ak_unique_asvs) <- host_names

saveRDS(ak_unique_asvs, "../intermediate_files/rds/ak_unique_asvs.rds")

```

## Identify Core ASVs

In this case, "core" refers to ASVs unique to the same host in both habitats

```{r}

core_unique_asvs <- list()

  
for (a_host in host_names) {
  
  core_asvs_by_host <- intersect(ro_unique_asvs[[a_host]]$ASV,
                                 ak_unique_asvs[[a_host]]$ASV)
  
  core_unique_list <- 
    ro_unique_asvs[[a_host]][
      ro_unique_asvs[[a_host]]$ASV %in% core_asvs_by_host, ]
  
  core_unique_asvs[[a_host]] <- core_unique_list$ASV
  
}
  

core_unique_names <- unlist(core_unique_asvs, use.names = F)


```

## Making Core Column

Previously I appended "_core" to each family name, so the names of objects and functions have "append" in them. Ignore that! We are now creating a "Core" column in the dataframe that will say "Yes" if an ASV is core and "No" if it is not. The ggplot alpha value will depend on this column.

```{r}

append_core <- function (a_host, list_of_dfs) {

  # append "_core" to a family name if the ASV is present in both communities
  list_of_dfs[[a_host]] <- mutate(list_of_dfs[[a_host]],
                        Core = ifelse(ASV %in% core_unique_names,
                                        paste("Yes"),
                                        paste("No")))

}

# run function and rename items
ro_core_appended <- lapply(host_names, 
                           append_core,
                           ro_unique_asvs)

names(ro_core_appended) <- host_names


# run function and rename items
ak_core_appended <- lapply(host_names, 
                           append_core,
                           ak_unique_asvs)

names(ak_core_appended) <- host_names



```

## rbind

```{r}

ro_unique_rbind <- do.call("rbind", ro_core_appended)

saveRDS(ro_unique_rbind, "../intermediate_files/rds/ro_unique_rbind.rds")
write.csv(ro_unique_rbind, "../intermediate_files/csv/ro_unique_rbind.csv")


ak_unique_rbind <- do.call("rbind", ak_core_appended)

saveRDS(ak_unique_rbind, "../intermediate_files/rds/ak_unique_rbind.rds")
write.csv(ak_unique_rbind, "../intermediate_files/csv/ak_unique_rbind.csv")


## Getting ASV counts for bar labels

sums_for_bars <- function (rbind_df) {
  rbind_sums <- rbind_df %>%
    group_by(Host) %>%
    summarise(count=length(ASV), abund=sum(rel_abun))
}


ro_unique_bar_sums <- sums_for_bars(ro_unique_rbind)

ak_unique_bar_sums <- sums_for_bars(ak_unique_rbind)


```

## Plotting

```{r}

source("../src/bar_plotting.R")

### arguments to the barplotting function:
# the rbinded dataframe of relative abundances for all hosts and soil
# the summarized data frame containing ASV counts to be used as bar labels
# the upper limit for the y axis (number)
# the vector of color names from bar_plotting.R
# the title for the plot (string)


ro_unique_bar <- core_barplot(ro_unique_rbind,
                                 ro_unique_bar_sums,
                                 8,
                                 colors_ten_families,
                                 "\nRemnant Forest")

ak_unique_bar <- core_barplot(ak_unique_rbind,
                                 ak_unique_bar_sums,
                                 8,
                                 colors_ten_families,
                                 "\nRestored Forest")

# arrange side by side
combined_unique_bar <- two_barplots(ro_unique_bar, ak_unique_bar)

# export
ggsave("../figures/core_unique_asv_barplot.png", width = 11.5, height = 7)



# arguments for the noncore (non-shaded) barplot are the same as above

ro_unique_bar_noncore <- noncore_barplot(ro_unique_rbind,
                                 ro_unique_bar_sums,
                                 8,
                                 colors_ten_families,
                                 "\nRemnant Forest")

ak_unique_bar_noncore <- noncore_barplot(ak_unique_rbind,
                                 ak_unique_bar_sums,
                                 8,
                                 colors_ten_families,
                                 "\nRestored Forest")

combined_unique_bar_noncore <- two_barplots(ro_unique_bar_noncore, ak_unique_bar_noncore)

ggsave("../figures/unique_asv_barplot.png", width = 11.5, height = 7)


```


### 2x2 plot

```{r}

# don't want titles for the core plots

### arguments include:
# the rbinded dataframe of relative abundances for all hosts and soil
# the summarized data frame containing ASV counts to be used as bar labels
# the upper limit for the y axis (number)
# the vector of color names from bar_plotting.R

ro_unique_bar_notitle <- core_barplot_notitle(ro_unique_rbind,
                                              ro_unique_bar_sums,
                                              8,
                                              colors_ten_families)

ak_unique_bar_notitle <- core_barplot_notitle(ak_unique_rbind,
                                              ak_unique_bar_sums,
                                              8,
                                              colors_ten_families)



# first arrange side by side plots for shared and unique ASVs to get
# common legends for both (different families are present in each grouping)

both_unique <- ggarrange(ro_unique_bar_noncore,
                         ak_unique_bar_noncore, 
                         labels=c("A)", "B)"), 
                         hjust=-0.5, 
                         vjust = 5.5, 
                         common.legend = TRUE, 
                         legend = "right")

both_unique_core <- ggarrange(ro_unique_bar_notitle,
                              ak_unique_bar_notitle,
                              labels=c("C)", "D)"),
                              hjust=-0.5,
                              vjust = 1,
                              common.legend = TRUE,
                              legend = "right")



# combine the pre-arranged side by side plots to make a 2x2 plot
all_unique_barplots <- ggarrange(both_unique,
                                 both_unique_core,
                                 ncol=1,
                                 nrow=2,
                                 heights = c(1.15, 1)) +
                  theme(plot.margin = margin(0,0,0.5,0.5, "cm"))

all_unique_barplots <- annotate_figure(all_unique_barplots,
                                left = text_grob("Relative Abundance (%)",
                                                 rot = 90,
                                                 size = 16,
                                                 hjust = 0.2),
                                bottom = text_grob("Host", size = 16, hjust = 2.7, vjust=-1))

ggsave("../figures/2x2_unique_barplots.png", width=11, height=12)


```





