---
title: "Soil + Host Barplots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}

library(tibble)
library(dplyr)
library(ggplot2)

```

### Read in necessary files

```{r}

# count data frames from unrarefied data
ro_barplot_df <- readRDS("../intermediate_files/rds/ro_barplot_df.rds")

ak_barplot_df <- readRDS("../intermediate_files/rds/ak_barplot_df.rds")

# taxonomy table from full unrarefied physeq object 
full_tax_table_df <- readRDS("../intermediate_files/rds/full_tax_table_df.rds")

# a vector of host names + soil
host_names <- readRDS("../intermediate_files/rds/host_names.rds")

```



```{r}

ro_soil_and_hosts <- list()


host_names_no_soil <- host_names[1:6]


# makes a data frame per host including ASVs, counts, relative abundance, family, and host names
get_soil_host_asvs <- function (a_host, barplot_df, taxonomy_df) {
  
  # for each host, subset ASVs that are shared between a host and soil
  soil_host_subset <- barplot_df[barplot_df[a_host] > 0 &
                                   barplot_df["Soil"] > 0, ]
  
  # make asv rownames into a column
  soil_host_subset <- rownames_to_column(soil_host_subset, "ASV")
  # make a new df with asv names and host sample asv counts as columns
  soil_host_subset <- select(soil_host_subset, ASV, a_host)
  # take sum of all unique host sample sequences
  total_host_sum <- sum(barplot_df[[a_host]])
  # make new relative abundance column by dividing unique count by total
  soil_host_subset <- mutate(soil_host_subset, rel_abun=((soil_host_subset[[a_host]] / total_host_sum) * 100))
  # make new column taking family from tax_table where asv names match in both dfs
  soil_host_subset$Family <- full_tax_table_df$Family[match(soil_host_subset$ASV, full_tax_table_df$ASV)]
  # make new column with host name filling all cells
  soil_host_subset$Host <- rep(a_host, nrow(soil_host_subset))
  # make new column with ASV total per host filling all cells
  soil_host_subset$Total <- rep(sum(soil_host_subset[[a_host]]), nrow(soil_host_subset))
  # change host colname for future rbind
  colnames(soil_host_subset)[2] <- "Count"

  return(soil_host_subset)
  
}


# run function

ro_soil_host_asvs <- lapply(host_names_no_soil, 
                         get_soil_host_asvs, 
                         ro_barplot_df, 
                         full_tax_table_df)

names(ro_soil_host_asvs) <- host_names_no_soil


ak_soil_host_asvs <- lapply(host_names_no_soil, 
                         get_soil_host_asvs, 
                         ak_barplot_df, 
                         full_tax_table_df)

names(ak_soil_host_asvs) <- host_names_no_soil

```

## Identify Core ASVs

In this case, "core" refers to ASVs unique to the same host in both habitats

```{r}

core_soil_host_asvs <- list()

  
for (a_host in host_names) {
  
  core_asvs_by_host <- intersect(ro_soil_host_asvs[[a_host]]$ASV,
                                 ak_soil_host_asvs[[a_host]]$ASV)
  
  core_soil_host_list <- 
    ro_soil_host_asvs[[a_host]][
      ro_soil_host_asvs[[a_host]]$ASV %in% core_asvs_by_host, ]
  
  core_soil_host_asvs[[a_host]] <- core_soil_host_list$ASV
  
}
  

core_soil_host_names <- unique(unlist(core_soil_host_asvs, use.names = F))

saveRDS(core_soil_host_names, "../intermediate_files/rds/core_soil_host_names.rds")


```

## Making Core Column

Previously I appended "_core" to each family name, so the names of objects and functions have "append" in them. Ignore that! We are now creating a "Core" column in the dataframe that will say "Yes" if an ASV is core and "No" if it is not. The ggplot alpha value will depend on this column.


```{r}

append_core_soil_host <- function (a_host, list_of_dfs) {

  # append "_core" to a family name if the ASV is present in both communities
  list_of_dfs[[a_host]] <- mutate(list_of_dfs[[a_host]],
                        Core = ifelse(ASV %in% core_soil_host_names,
                                        paste("Yes"),
                                        paste("No")))

}


ro_sh_core_appended <- lapply(host_names_no_soil, 
                           append_core_soil_host,
                           ro_soil_host_asvs)

names(ro_sh_core_appended) <- host_names_no_soil



ak_sh_core_appended <- lapply(host_names_no_soil, 
                           append_core_soil_host,
                           ak_soil_host_asvs)

names(ak_sh_core_appended) <- host_names_no_soil

```

## rbind

```{r}

ro_soil_host_rbind <- do.call("rbind", ro_sh_core_appended)

saveRDS(ro_soil_host_rbind, "../intermediate_files/rds/ro_soil_host_rbind.rds")
write.csv(ro_soil_host_rbind, "../intermediate_files/csv/ro_soil_host_rbind.csv")


ak_soil_host_rbind <- do.call("rbind", ak_sh_core_appended)

saveRDS(ak_soil_host_rbind, "../intermediate_files/rds/ak_soil_host_rbind.rds")
write.csv(ak_soil_host_rbind, "../intermediate_files/csv/ak_soil_host_rbind.csv")



## Getting ASV counts for bar labels

sums_for_bars <- function (rbind_df) {
  rbind_sums <- rbind_df %>%
    group_by(Host) %>%
    summarise(count=length(ASV), abund=sum(rel_abun))
}


ro_soil_host_bar_sums <- sums_for_bars(ro_soil_host_rbind)

ak_soil_host_bar_sums <- sums_for_bars(ak_soil_host_rbind)


```

## Plotting

```{r}

source("../src/bar_plotting.R")

### arguments to the barplotting function:
# the rbinded dataframe of relative abundances for all hosts and soil
# the summarized data frame containing ASV counts to be used as bar labels
# the upper limit for the y axis (number)
# the vector of color names from bar_plotting.R
# the title for the plot (string)


ro_soil_host_bar <- core_barplot(ro_soil_host_rbind,
                                 ro_soil_host_bar_sums,
                                 100,
                                 colors_ten_families,
                                 "\nRemnant Forest")

ak_soil_host_bar <- core_barplot(ak_soil_host_rbind,
                                 ak_soil_host_bar_sums,
                                 100,
                                 colors_ten_families,
                                 "\nRestored Forest")

combined_soil_host_bar <- two_barplots(ro_soil_host_bar, ak_soil_host_bar)


ggsave("../figures/core_soil_host_barplot.png", width = 11, height = 7)


# arguments are the same for noncore function

ro_soho_bar_noncore <- noncore_barplot(ro_soil_host_rbind,
                                 ro_soil_host_bar_sums,
                                 100,
                                 colors_ten_families,
                                 "\nRemnant Forest")

ak_soho_bar_noncore <- noncore_barplot(ak_soil_host_rbind,
                                 ak_soil_host_bar_sums,
                                 100,
                                 colors_ten_families,
                                 "\nRestored Forest")

combined_soho_bar_noncore <- two_barplots(ro_soho_bar_noncore, ak_soho_bar_noncore)

ggsave("../figures/soil_host_barplot.png", width = 11, height = 7)


```



