---
title: "Generalist Barplots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}

library(tibble)
library(dplyr)
library(ggplot2)

```

### Read in necessary files

```{r}

# count data frames from unrarefied data
ro_barplot_df <- readRDS("../intermediate_files/rds/ro_barplot_df.rds")

ak_barplot_df <- readRDS("../intermediate_files/rds/ak_barplot_df.rds")

# taxonomy table from full unrarefied physeq object 
full_tax_table_df <- readRDS("../intermediate_files/rds/full_tax_table_df.rds")

# a vector of host names + soil
host_names <- readRDS("../intermediate_files/rds/host_names.rds")

```

## Prepare data frame for ASVs shared between all hosts and soil

```{r}

get_shared_asvs <- function (a_host, barplot_df, taxonomy_df) {
  
  # subset host specific ASVs by host. Counts for all other hosts should be 0, 
  # so the rowsum == the host count
  shared_subset <- barplot_df[apply(barplot_df, 1, function(row) all(row!=0)),]
  
  # make asv rownames into a column
  shared_subset <- rownames_to_column(shared_subset, "ASV")
  
  # make a new df with asv names and host sample asv counts as columns
  shared_subset <- select(shared_subset, ASV, a_host)
  
  # take sum of all shared host sample sequences
  total_host_sum <- sum(barplot_df[[a_host]])
  
  # make new relative abundance column by dividing shared count by total
  shared_subset <- mutate(shared_subset, rel_abun=((shared_subset[[a_host]] / total_host_sum) * 100))
  
  # make new column taking family from tax_table 
  # where asv names match in both dfs
  shared_subset$Family <- 
    taxonomy_df$Family[match(shared_subset$ASV, taxonomy_df$ASV)]
  
  # make new column with host name filling all cells
  shared_subset$Host <- rep(a_host, nrow(shared_subset))
  
  # make new column with ASV total per host filling all cells
  shared_subset$Total <- rep(sum(shared_subset[[a_host]]), nrow(shared_subset))
  
  # change host colname for future rbind
  colnames(shared_subset)[2] <- "Count"
  
  return(shared_subset)

}


# run function

ro_shared_asvs <- lapply(host_names, 
                         get_shared_asvs, 
                         ro_barplot_df, 
                         full_tax_table_df)

names(ro_shared_asvs) <- host_names

saveRDS(ro_shared_asvs, "../intermediate_files/rds/ro_shared_asvs.rds")



ak_shared_asvs <- lapply(host_names, 
                         get_shared_asvs, 
                         ak_barplot_df, 
                         full_tax_table_df)

names(ak_shared_asvs) <- host_names

saveRDS(ak_shared_asvs, "../intermediate_files/rds/ak_shared_asvs.rds")


```

## Identify Core ASVs

In this case, "core" refers to ASVs shared to the same host in both habitats

```{r}

ro_shared_names <- ro_shared_asvs[["A. koa"]]$ASV

ak_shared_names <- ak_shared_asvs[["A. koa"]]$ASV

core_shared_names <- intersect(ro_shared_names, ak_shared_names)


```

## Making Core Column

Previously I appended "_core" to each family name, so the names of objects and functions have "append" in them. Ignore that! We are now creating a "Core" column in the dataframe that will say "Yes" if an ASV is core and "No" if it is not. The ggplot alpha value will depend on this column.

```{r}

append_core_shared <- function (a_host, list_of_dfs) {

  # append "_core" to a family name if the ASV is present in both communities
  list_of_dfs[[a_host]] <- mutate(list_of_dfs[[a_host]],
                        "In Both Habitats" = ifelse(ASV %in% core_shared_names,
                                        paste("Yes"),
                                        paste("No")))

}


ro_shared_core_appended <- lapply(host_names, 
                           append_core_shared,
                           ro_shared_asvs)

names(ro_shared_core_appended) <- host_names



ak_shared_core_appended <- lapply(host_names, 
                           append_core_shared,
                           ak_shared_asvs)

names(ak_shared_core_appended) <- host_names

```

## rbind

```{r}

ro_shared_rbind <- do.call("rbind", ro_shared_core_appended)

saveRDS(ro_shared_rbind, "../intermediate_files/rds/ro_shared_rbind.rds")
write.csv(ro_shared_rbind, "../intermediate_files/csv/ro_shared_rbind.csv")


ak_shared_rbind <- do.call("rbind", ak_shared_core_appended)

saveRDS(ak_shared_rbind, "../intermediate_files/rds/ak_shared_rbind.rds")
write.csv(ak_shared_rbind, "../intermediate_files/csv/ak_shared_rbind.csv")



```

## Plotting

```{r}

source("../src/bar_plotting.R")

### arguments to the barplotting function:
# the rbinded dataframe of relative abundances for all hosts and soil
# the upper limit for the y axis (number)
# the vector of color names from bar_plotting.R
# the title for the plot (string)

ro_shared_bar <- make_barplot(ro_shared_rbind,
                                 100,
                                 colors_ten_families,
                                 "\nRemnant Forest") 


ak_shared_bar <- make_barplot(ak_shared_rbind,
                                 100,
                                 colors_ten_families,
                                 "\nRestored Forest") 


combined_shared_bar <- two_barplots(ro_shared_bar, ak_shared_bar)


ggsave("../figures/core_shared_barplot3.png", width = 11, height = 7)


# arguments are the same for noncore function
# 
# ro_shared_bar_noncore <- noncore_barplot(ro_shared_rbind,
#                                  100,
#                                  colors_ten_families,
#                                  "\nRemnant Forest") 
# 
# ak_shared_bar_noncore <- noncore_barplot(ak_shared_rbind,
#                                  100,
#                                  colors_ten_families,
#                                  "\nRestored Forest")
# 
# 
# combined_shared_bar_noncore <- two_barplots(ro_shared_bar_noncore, ak_shared_bar_noncore)
# 
# # Export
# ggsave("../figures/shared_barplot.png", width = 11, height = 7)


```






